---
title: "SSR tolerence set by Chai-Squared distribution"
output:
  pdf_document: default
  html_document: default
  
  #instructions
  #line 23 model, history, compresultsが保存されているfolder nameを入れる
  #line 35 namually choose maximun number of accepted parameters
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
#import libs
```{r}
library(ggplot2)
library(car)
```
# setup
- Set the directory where the model file and compResult is
```{r}
foldername='//Volumes/desktop/'

tempFname=strsplit(foldername, "/")
Fname=tempFname[[1]][length(tempFname[[1]])]

iterationSetting=as.numeric(read.csv(paste0(foldername,"/model/iterationSetting.csv"),header = FALSE)$V1)
```
- p value
- to use L-shaped after chi square (more strict acceptance criterion)
```{r}
pvalTolerence=0.05
strictSelection=TRUE #set to TRUE if want to apply L-shape method
manuallyChosenAcceptNumParam=NA #set to NA if want to automatically choose it by Chai-square
```
#import and process data
```{r}
sortIndex=read.table(paste0(foldername,'/compResult/sortIndex.csv'), sep=',', fill = TRUE)
iterNumber=iterationSetting[1]+iterationSetting[2]-1
R_df=read.table(paste0(foldername,'/compResult/R_',iterNumber,'.csv'), sep=',')
Y_df=read.table(paste0(foldername,'/compResult/Y_',iterNumber,'.csv'), sep=',')

observations_df=read.table(paste0(foldername,'/model/observations.txt'), sep=',')
observations_df=observations_df[!is.na(observations_df$V1)&!is.na(observations_df$V2)&!is.na(observations_df$V3),]
observations_df=observations_df[seq(1,dim(Y_df)[2]),]
y_star=c()
for(text_in in observations_df$V3){
  y_star=c(y_star,eval(parse(text=(tolower(text_in)))))
}

obsTime=c()
for(text_in in observations_df$V1){
  obsTime=c(obsTime,eval(parse(text=(tolower(text_in)))))
}
observations_df$V1=obsTime

obsTime=unique(observations_df$V1)
#obsTime=as.numeric(read.table(paste0(foldername,'/compResult/ode_time.csv'), sep=','))


SSR_vec=as.numeric(R_df)
min_R=min(R_df, na.rm = TRUE)
min_index=which(R_df==min_R)[1]

sum((Y_df[min_index,]-y_star)^2)
residual_vec=as.numeric(Y_df[min_index,]-y_star)
mean(residual_vec)
sd(residual_vec)
ggplot(data.frame(residual=residual_vec), aes(residual))+geom_histogram(bins=10)
```

# conduct test
```{r}

qchisq(1-pvalTolerence, df=length(residual_vec)) *(sd(residual_vec))^2

accept_index=which(R_df<(qchisq(1-pvalTolerence, df=length(residual_vec)) *(sd(residual_vec))^2+min_R))

acceptMaxSSR=qchisq(1-pvalTolerence, df=length(residual_vec)) *(sd(residual_vec))^2+min_R
accept_vec=as.vector(SSR_vec<acceptMaxSSR)

numAccept=sum(accept_vec, na.rm = TRUE)

```
#Accepted maximum SSR is `r acceptMaxSSR` and `r numAccept` parameter sets were accepted by chai-squared method.


#plot SSR v.s. rank
```{r}
if(strictSelection){
  
if(!is.na(manuallyChosenAcceptNumParam)){
  sortedAcceptedSSR=sort(SSR_vec[accept_vec])[seq(1,manuallyChosenAcceptNumParam)]
}else{
  sortedAcceptedSSR=sort(SSR_vec[accept_vec])
}


trapizoido_area=c()
for(i in seq(1,length(sortedAcceptedSSR)-1))
  trapizoido_area=c(trapizoido_area, (sortedAcceptedSSR[1]+sortedAcceptedSSR[i])*i+(sortedAcceptedSSR[i+1]+sortedAcceptedSSR[length(sortedAcceptedSSR)])*(length(sortedAcceptedSSR)-i))

strictMaxAcceptSSR=sortedAcceptedSSR[which(trapizoido_area==min(trapizoido_area))]

accept_index=which(SSR_vec<strictMaxAcceptSSR)
accept_vec=as.vector(SSR_vec<strictMaxAcceptSSR)
numAccept=sum(accept_vec, na.rm = TRUE)
acceptMaxSSR=strictMaxAcceptSSR

}
 

```
#Accepted maximum SSR is `r strictMaxAcceptSSR` and `r numAccept` parameter sets were accepted.


```{r}
 plot_df=data.frame(SSR=SSR_vec, order=rank(SSR_vec), accept=accept_vec)

ggplot(plot_df, aes(x=order,y=SSR, colour=accept))+geom_point()+coord_cartesian(ylim=c(0,acceptMaxSSR*2))+geom_vline(xintercept = numAccept, color="grey")+ annotate(geom="text", x=numAccept, y=acceptMaxSSR*0.5, label=paste("Accepted: ",numAccept,"\n Accepted max SSR: ",formatC(acceptMaxSSR, format = "g", digits = 3)),angle = 90,
              color="black")+ annotate(geom="text", x=length(SSR_vec)*0.1, y=min_R*1.1, label=paste("min SSR: ",formatC(min_R, format = "g", digits = 3)),
              color="black")+labs(caption=paste(ifelse(strictSelection,ifelse(is.na(manuallyChosenAcceptNumParam),"Cutoff chosen by chai square and then applied L-shape",paste("Cutoff chosen by user to be ",manuallyChosenAcceptNumParam, "and then applied L-shape")),"Cutoff chosen by chai square"),"\n plot based on", foldername))
```
# Residual plots for the parameter set with minimum SSR
```{r}
#indexToShow=which(SSR_vec==max(SSR_vec[accept_vec]))
indexToShow=min_index
SSR_vec[indexToShow]

temp_residual_vec=as.numeric(Y_df[indexToShow,]-y_star)


min=resigual_df=data.frame(square_residual=temp_residual_vec^2,variableNames=observations_df$V2)
SSR_byVariable=aggregate(resigual_df$square_residual, by=list(variableNames=resigual_df$variableNames), FUN=sum)
SSR_byVariable$SSR=formatC(SSR_byVariable$x, format = "g", digits = 3)

ggplot(data.frame(residual=temp_residual_vec,prediction=as.numeric(Y_df[indexToShow,]),variableNames=observations_df$V2), aes(y=residual,x=prediction))+geom_point()+geom_hline(yintercept = 0)+geom_smooth()

ggplot(data.frame(residual=temp_residual_vec,prediction=as.numeric(Y_df[indexToShow,]),variableNames=observations_df$V2), aes(y=residual,x=prediction))+geom_point()+geom_hline(yintercept = 0)+geom_smooth()+facet_wrap(variableNames~., scales = "free_x")+
  geom_text(
    size    = 5,
    data    = SSR_byVariable,
    mapping = aes(x = Inf, y = Inf, label = paste0("SSR=",SSR)),
    hjust   = 1.05,
    vjust   = 1.5
  )


ggplot(data.frame(residual=temp_residual_vec,observations=y_star,variableNames=observations_df$V2), aes(y=residual,x=observations))+geom_point()+geom_hline(yintercept = 0)+geom_smooth()

ggplot(data.frame(residual=temp_residual_vec,observations=y_star,variableNames=observations_df$V2), aes(y=residual,x=observations))+geom_point()+geom_hline(yintercept = 0)+geom_smooth()+facet_wrap(variableNames~., scales = "free_x")+
  geom_text(
    size    = 5,
    data    = SSR_byVariable,
    mapping = aes(x = Inf, y = Inf, label = paste0("SSR=",SSR)),
    hjust   = 1.05,
    vjust   = 1.5
  )


qqPlot(temp_residual_vec)


```

# Residual plots for the parameter set with maximum accepted SSR

```{r}
indexToShow=which(SSR_vec==max(SSR_vec[accept_vec]))
SSR_vec[indexToShow]

temp_residual_vec=as.numeric(Y_df[indexToShow,]-y_star)


min=resigual_df=data.frame(square_residual=temp_residual_vec^2,variableNames=observations_df$V2)
SSR_byVariable=aggregate(resigual_df$square_residual, by=list(variableNames=resigual_df$variableNames), FUN=sum)
SSR_byVariable$SSR=formatC(SSR_byVariable$x, format = "g", digits = 3)

ggplot(data.frame(residual=temp_residual_vec,prediction=as.numeric(Y_df[indexToShow,]),variableNames=observations_df$V2), aes(y=residual,x=prediction))+geom_point()+geom_hline(yintercept = 0)+geom_smooth()

ggplot(data.frame(residual=temp_residual_vec,prediction=as.numeric(Y_df[indexToShow,]),variableNames=observations_df$V2), aes(y=residual,x=prediction))+geom_point()+geom_hline(yintercept = 0)+geom_smooth()+facet_wrap(variableNames~., scales = "free_x")+
  geom_text(
    size    = 5,
    data    = SSR_byVariable,
    mapping = aes(x = Inf, y = Inf, label = paste0("SSR=",SSR)),
    hjust   = 1.05,
    vjust   = 1.5
  )


ggplot(data.frame(residual=temp_residual_vec,observations=y_star,variableNames=observations_df$V2), aes(y=residual,x=observations))+geom_point()+geom_hline(yintercept = 0)+geom_smooth()

ggplot(data.frame(residual=temp_residual_vec,observations=y_star,variableNames=observations_df$V2), aes(y=residual,x=observations))+geom_point()+geom_hline(yintercept = 0)+geom_smooth()+facet_wrap(variableNames~., scales = "free_x")+
  geom_text(
    size    = 5,
    data    = SSR_byVariable,
    mapping = aes(x = Inf, y = Inf, label = paste0("SSR=",SSR)),
    hjust   = 1.05,
    vjust   = 1.5
  )


qqPlot(temp_residual_vec)


```

# plot time course simulation
```{r}
numSamples=dim(Y_df)[1]
numObs=dim(Y_df)[2]
observations_df=observations_df[1:numObs,]

timeCoursePlot_df=data.frame(time=rep(observations_df$V1, numSamples), sim=as.vector(t(Y_df)), variable=rep(observations_df$V2, numSamples),sampleIndex=as.factor(rep(seq(1,numSamples), each = numObs)))

timeCoursePlot_df$accept=with(timeCoursePlot_df,sampleIndex%in%accept_index)

ggplot(timeCoursePlot_df, aes(x=time,y=sim, group=sampleIndex,colour=accept))+geom_line(alpha=0.2)+geom_point(size=2,shape=4)+geom_point(data=data.frame(time=observations_df$V1, obs=y_star, variable=observations_df$V2,accept=NA), aes(x=time,y=obs,group=variable))+facet_wrap(variable~., scales='free_y')+coord_cartesian(ylim = c(min(y_star)-(max(y_star)-min(y_star))*0.1,max(y_star)+(max(y_star)-min(y_star))*0.1))+labs(caption = paste("plot based on", foldername))


ggplot(timeCoursePlot_df, aes(x=time,y=sim, group=sampleIndex,colour=accept))+geom_line(alpha=0.2)+geom_point(size=2,shape=4)+geom_point(data=data.frame(time=observations_df$V1, obs=y_star, variable=observations_df$V2,accept=NA), aes(x=time,y=obs,group=variable))+facet_wrap(variable~., scales='free_y')+labs(caption = paste("plot based on", foldername))
```

```{r}
numSamples=dim(Y_df)[1]
numObs=dim(Y_df)[2]
observations_df=observations_df[1:numObs,]

timeCoursePlot_df=data.frame(time=rep(observations_df$V1, numSamples), sim=as.vector(t(Y_df)), variable=rep(observations_df$V2, numSamples),sampleIndex=as.factor(rep(seq(1,numSamples), each = numObs)))

timeCoursePlot_df$accept=with(timeCoursePlot_df,sampleIndex%in%accept_index)

ggplot(subset(timeCoursePlot_df,accept==TRUE), aes(x=time,y=sim, group=sampleIndex))+geom_line(alpha=0.2)+geom_point(data=data.frame(time=observations_df$V1, obs=y_star, variable=observations_df$V2), aes(x=time,y=obs,group=variable), colour="red")+facet_wrap(variable~., scales='free_y')+coord_cartesian(ylim = c(min(y_star)-(max(y_star)-min(y_star))*0.1,max(y_star)+(max(y_star)-min(y_star))*0.1))+labs(caption = paste("plot based on", foldername))


ggplot(subset(timeCoursePlot_df,accept==TRUE), aes(x=time,y=sim, group=sampleIndex))+geom_line(alpha=0.2)+geom_point(data=data.frame(time=observations_df$V1, obs=y_star, variable=observations_df$V2), aes(x=time,y=obs,group=variable), colour="red")+facet_wrap(variable~., scales='free_y')+labs(caption = paste("plot based on", foldername))
```



```{r}

InitialParameter=read.table(paste0(foldername,'/compResult/Parameter_0.csv'), sep=',')
ParameterNames=as.vector(read.table(paste0(foldername,'/model/parameterNames.txt'), sep=',')$V1)
names(InitialParameter)=ParameterNames

```

```{r}

FinalParameter=read.table(paste0(foldername,'/compResult/Parameter_',iterNumber,'.csv'), sep=',')
names(FinalParameter)=ParameterNames

```


```{r}

freeParameterNames=c()
freeParaValues=data.frame()
data_forTable=data.frame()
for(pName in ParameterNames){
  if(min(InitialParameter[,pName])!=max(InitialParameter[,pName])){
    freeParameterNames=c(freeParameterNames,pName)
    freeParaValues=rbind(freeParaValues, data.frame(Name=pName, Value=InitialParameter[,pName], Kind="Initial"), data.frame(Name=pName, Value=FinalParameter[accept_vec,pName], Kind="Final"))
    data_forTable=rbind(data_forTable, data.frame(Name=pName, minimum=formatC(quantile(FinalParameter[accept_vec,pName], probs = c(0),na.rm = TRUE), format = "g", digits = 3), percentile_25=formatC(quantile(FinalParameter[accept_vec,pName], probs = c(0.25),na.rm = TRUE), format = "g", digits = 3), median=formatC(quantile(FinalParameter[accept_vec,pName], probs = c(0.5),na.rm = TRUE), format = "g", digits = 3), percentile_75=formatC(quantile(FinalParameter[accept_vec,pName], probs = c(0.75),na.rm = TRUE), format = "g", digits = 3), maximum=formatC(quantile(FinalParameter[accept_vec,pName], probs = c(1),na.rm = TRUE), format = "g", digits = 3)))
  }
  
}

freeParaValues$Kind=factor(as.character(freeParaValues$Kind),level=c("Initial","Final"))
```


# Violin plot with median and interquartile range (in Log scale)
```{r}
p<-ggplot(freeParaValues,aes(x=Kind,y=Value))+facet_wrap(Name~., scales = "free")+scale_y_continuous(trans = 'log10')

p+geom_violin(trim=T,fill="#999999",linetype="blank",alpha=I(1/2))+
 stat_summary(geom="pointrange",fun.y = median, fun.ymin = function(x) quantile(x,probs=0.25), fun.ymax = function(x) quantile(x,probs=0.75), size=0.5,alpha=.5)+
  stat_summary(geom="line",fun.y =  function(x) quantile(x,probs=0), aes(group=1),size=0.5,alpha=.3,linetype=2)+
  stat_summary(geom="line",fun.y =  function(x) quantile(x,probs=1), aes(group=1),size=0.5,alpha=.3,linetype=2)+
  stat_summary(geom="line",fun.y =  function(x) quantile(x,probs=0.25), aes(group=1),size=0.5,alpha=.3)+
  stat_summary(geom="line",fun.y =  function(x) quantile(x,probs=0.75), aes(group=1),size=0.5,alpha=.3)+
   theme(legend.position="none")+labs(caption = paste("plot based on", foldername))
```

# Violin plot with median and interquartile range
```{r}
p<-ggplot(freeParaValues,aes(x=Kind,y=Value))+facet_wrap(Name~., scales = "free")


p+geom_violin(trim=T,fill="#999999",linetype="blank",alpha=I(1/2))+
 stat_summary(geom="pointrange",fun.y = median, fun.ymin = function(x) quantile(x,probs=0.25), fun.ymax = function(x) quantile(x,probs=0.75), size=0.5,alpha=.5)+
  stat_summary(geom="line",fun.y =  function(x) quantile(x,probs=0), aes(group=1),size=0.5,alpha=.3,linetype=2)+
  stat_summary(geom="line",fun.y =  function(x) quantile(x,probs=1), aes(group=1),size=0.5,alpha=.3,linetype=2)+
  stat_summary(geom="line",fun.y =  function(x) quantile(x,probs=0.25), aes(group=1),size=0.5,alpha=.3)+
  stat_summary(geom="line",fun.y =  function(x) quantile(x,probs=0.75), aes(group=1),size=0.5,alpha=.3)+
   theme(legend.position="none")+labs(caption = paste("plot based on", foldername))
```
# table of summary statistics
```{r}
rownames(data_forTable)=NULL
(data_forTable)
```

#Bootstrap analysis visualisation (comment out when not running bootstrap analysis)

```{r}
bootstrapIter=iterationSetting[1]+iterationSetting[2]+iterationSetting[3]-1
postBootstrapPara=(read.csv(file=paste0(foldername,"/compResult/Parameter_",bootstrapIter,".csv"),header=FALSE))
usedForBootstrap=!is.na(as.list(read.csv(file=paste0(foldername,"/compResult/R_",bootstrapIter,".csv"),header=FALSE)))

names(postBootstrapPara)=ParameterNames
```

```{r}

freeParaBootstrapValues=data.frame()
for(pName in freeParameterNames){
  freeParaBootstrapValues=rbind(freeParaBootstrapValues, data.frame(Name=pName, Value=postBootstrapPara[usedForBootstrap,pName], Kind="Bootstrap"))
}

ggplot(freeParaBootstrapValues, aes(x=Value))+geom_histogram()+facet_wrap(.~Name,scales="free")
ggplot(freeParaBootstrapValues, aes(x=log10(Value)))+geom_histogram()+facet_wrap(.~Name,scales="free")


```
```{r}
data_forTable=data.frame()

for(pName in freeParameterNames){
data_forTable=rbind(data_forTable, data.frame(Name=pName, minimum=formatC(quantile(FinalParameter[accept_vec,pName], probs = c(0),na.rm = TRUE), format = "g", digits = 3), percentile_25=formatC(quantile(FinalParameter[accept_vec,pName], probs = c(0.25),na.rm = TRUE), format = "g", digits = 3), median=formatC(quantile(FinalParameter[accept_vec,pName], probs = c(0.5),na.rm = TRUE), format = "g", digits = 3), percentile_75=formatC(quantile(FinalParameter[accept_vec,pName], probs = c(0.75),na.rm = TRUE), format = "g", digits = 3), maximum=formatC(quantile(FinalParameter[accept_vec,pName], probs = c(1),na.rm = TRUE), format = "g", digits = 3), RSE=formatC(sd(postBootstrapPara[usedForBootstrap,pName], na.rm = TRUE)/mean(postBootstrapPara[usedForBootstrap,pName], na.rm = TRUE)*100, format = "g", digits = 3)))
}

```

```{r}
rownames(data_forTable)=NULL
(data_forTable)
```
